### 2023-09-07 Stats Rethinking/Ch3/Notes.qmd
Took notes through 3.2

---
title: "Notes"
format: html
---

# Libraries

```{r}
library(rethinking)
```

# 3.1 - Sampling from a grid-approximate posterior

```{r}
#Add the prior dist which we set to a uniform dist
p_grid <- seq(from=0, to=1, length.out=1000)
prob_p <- rep(1,1000)

#Add the "data" probability
prob_data <- dbinom(6, size=9, prob=p_grid)

#Multiply the two to get the posterior dist
posterior <- prob_data * prob_p
posterior <- posterior / sum(posterior)
```

```{r}
#Grab (or sample) values from our posterior and graph them below
#1e4 = 10,000
samples <- sample(p_grid, prob=posterior, size = 1e4, replace = TRUE)
```

```{r}
plot(samples)
```

```{r}
dens(samples)
```

# 3.2 - Sampling to Summarize

We can use bounds to find information about our parameters

```{r}
#What is the probability that the proporiotn of water is less than 0.5?
sum(posterior[p_grid < 0.5])
```