### 2023-09-08 Stats Rethinking/Ch3/Dr_Parson_ch_3_problems.Rmd
Create Dr_Parson_ch_3_problems.Rmd
Share my work on Chapter 3 practice problems.

---
title: "Chapter 3 practice problems"
author: "James Parson"
date: '2023-09-07'
output: html_document
---

```{r}
library(tidyverse)
library(rethinking)
```

# Sampling from prior

```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 / 1000 , 1000 )
samples <- sample( p_grid , prob=prior , size=1e4 , replace=TRUE )
```

```{r}
simulated_prior <- rbinom(n = 1e4, size = 15, prob = samples)

ggplot() +
  geom_histogram(mapping = aes(x = simulated_prior), bins = 100)
```

# Set-up for E problems

```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
```

# 3E1

```{r}
sum(samples <= 0.2) / 1e4
```

#3e2

```{r}
sum(samples >= 0.8) / 1e4
```

#3e3

```{r}
sum(samples > 0.2 & samples < 0.8) / 1e4
```

#3e4
```{r}
quantile(samples, 0.2)
```

#3e5
```{r}
quantile(samples, 0.8)
```

#3e6

```{r}
HPDI(samples, prob = 0.66)
```

```{r}
0.7737738 - 0.5085085
```

#3e7

```{r}
quantile(samples, c(0.5 - 0.66/2, 0.5 + 0.66/2))
```

```{r}
0.7697698 - 0.5025025
```

A slightly wider interval!

# 3m1

```{r}
N <- 1e3
grid <- seq(from = 0, to = 1, length.out = N)
prior <- rep(1, N)
likelihood <- dbinom(x = 8, size = 15, prob = grid)
posterior.raw <- prior * likelihood
posterior <- posterior.raw / sum(posterior.raw)

ggplot() +
  geom_line(mapping = aes(x = grid, y = posterior * N))
```
I scaled the probabilities in the posterior by N for the graph to match the vertical scale of the density below.

# 3m2

```{r}
samples <- sample(grid, size = 1e4, replace = TRUE, prob = posterior)

ggplot() +
  geom_density(mapping = aes(x = samples))
```

```{r}
HPDI(samples, prob = 0.9)
```

```{r}
water <- rbinom(1e4, 15, prob = samples)

ggplot() +
  geom_histogram(mapping = aes(x = water), bins = 100)
```

```{r}
sum(water == 8) / 1e4
```

Comparing Bayesian result with maximum-likelihood model.

```{r}
dbinom(8, 15, prob = 8/15)
```

Heavier tails in the predictive distribution from the Bayesian model.

```{r}
max_likelihood <- rbinom(1e4, size = 15, prob = 8 / 15)

ggplot() +
  geom_freqpoly(mapping = aes(x = water), bins = 100) +
  geom_freqpoly(mapping = aes(x = max_likelihood), bins = 100, color = "red")
```

#3m4

```{r}
water2 <- rbinom(1e4, size = 9, prob = samples)
```

```{r}
sum(water2 == 6) / 1e4
```

#3m5

```{r}
N <- 1e3
grid <- seq(from = 0, to = 1, length.out = N)
prior <- as.numeric(grid >= 0.5)
likelihood <- dbinom(x = 8, size = 15, prob = grid)
posterior.raw <- prior * likelihood
posterior <- posterior.raw / sum(posterior.raw)

ggplot() +
  geom_line(mapping = aes(x = grid, y = posterior * N))
```

```{r}
samples <- sample(grid, size = 1e5, replace = TRUE, prob = posterior)

ggplot() +
  geom_density(mapping = aes(x = samples))
```

```{r}
HPDI(samples, prob = 0.9)
```

```{r}
water <- rbinom(1e4, 15, prob = samples)

ggplot() +
  geom_histogram(mapping = aes(x = water), bins = 100)
```

```{r}
sum(water == 8) / 1e4
```

Comparing again with maximum-likelihood model.

```{r}
dbinom(8, 15, prob = 8/15)
```

```{r}
max_likelihood <- rbinom(1e4, size = 15, prob = 8 / 15)

ggplot() +
  geom_freqpoly(mapping = aes(x = water), bins = 100) +
  geom_freqpoly(mapping = aes(x = max_likelihood), bins = 100, color = "red")
```

```{r}
water2 <- rbinom(1e4, size = 9, prob = samples)
```

```{r}
sum(water2 == 6) / 1e4
```

# 3m6

Not sure what he is getting at! I'll have to come back to this one.

# 3h1

```{r}
data(homeworkch3)
```

```{r}
N <- 1e3
prior <- rep(1, N)
grid <- seq(from = 0, to = 1, length.out = N)
likelihood <- dbinom(x = 111, size = 200, prob = grid)
posterior.raw = prior * likelihood
posterior = posterior.raw / sum(posterior.raw)

ggplot() +
  geom_line(mapping = aes(x = grid, y = posterior * N))

```
```{r}
grid[which.max(posterior)]
```

```{r}
111/200
```

As expected! Maximum posterior probability (flat prior) matches maximum likelihood...

#3h2

```{r}
samples <- sample(grid, 1e4, replace = TRUE, prob = posterior)
```

```{r}
HPDI(samples, prob = c(0.5, 0.89, 0.97))
```

#3h3

```{r}
simulated <- rbinom(1e4, size = 200, prob = samples)
```

```{r}
ggplot() +
  geom_density(mapping = aes(x = simulated))
```
#3h4

```{r}
simulated_100 <- rbinom(1e4, size = 100, prob = samples)
```

```{r}
ggplot() +
  geom_density(mapping = aes(x = simulated_100))
```

```{r}
sum(birth1) / 100
```

Seeing at most 51 male births out of 100 is somewhat unusual but not terribly strange.

```{r}
sum(simulated_100 <= 51) / 1e4
```

#3h5

```{r}
births <- tibble(birth1 = birth1, birth2 = birth2)
```

```{r}
births %>% filter(birth1 == 0, birth2 == 1) %>% count()
```
```{r}
births %>% filter(birth1 == 0) %>% count()
```
```{r}
simulate_girl_first <- rbinom(n = 1e4, size = 49, prob = samples)
```

```{r}
ggplot() +
  geom_density(mapping = aes(x = simulate_girl_first)) +
  geom_vline(mapping = aes(xintercept = 39), color = "red")
```

```{r}
sum(simulate_girl_first >= 39) / 1e4
```

Seeing so many male second births given female first birth is very surprising given the model. Sexes of first and second birth do not appear to be independent.