### 2023-09-14 Stats Rethinking/Ch4/Practice Problems.qmd
Completed Easy Practice Problems

---
title: "Practice Problems"
format: html
---

# Load Libraries

```{r}
library(rethinking)
library(tidyverse)
```

# 4E1

In the model definition, which is the likelihood?

$$
y_i\sim Normal(\mu,\sigma)\\
\mu\sim Normal(0,10)\\
\sigma\sim Exponential(1)
$$

The likelihood is, $y_i\sim Normal(\mu,\sigma)$

# 4E2

In the model definition above, how many parameters are in the posterior distribution?

There are two parameters, $\mu$ and $\sigma$.

# 4E3

Using the model definition above, write down the appropriate form of Bayes' theorem that includes the proper likelihood and priors

$$
Pr(\mu,\sigma|y_i)=\frac{\prod_i Normal(y_i|\mu,\sigma)Normal(\mu|0,10)Exponential(\sigma|1)}{\int\int\prod_i Normal(y_i|\mu,\sigma)Normal(\mu|0,10)Exponential(\sigma|1)d\mu d\sigma}
$$

(Used formatting of Bayes on page 84)

# 4E4

In the model definition, which line is the linear model?

$$
y_i\sim Normal(\mu, \sigma)\\
\mu_i=\alpha+\beta x_i\\
\alpha\sim Normal(0,10)\\
\beta\sim Normal(0,1)\\
\sigma\sim Exponential(2)
$$

In this model, the $\mu_i=\alpha+\beta x_i$ is the linear model.

# 4E5

In the model definition for 4E4, how many parameters are in the posterior distribution?

There are 3 parameters in the posterior dist.

# 4M1

# 4M2

# 4M3

# 4M4

# 4M5

# 4M6

# 4M7

# 4M8

# 4H1

# 4H2

# 4H3

# 4H4

# 4H5

# 4H6

# 4H7

# 4H8