### 2023-09-14 Stats Rethinking/Ch4/Notes.qmd

---
title: "Notes"
format: html
---

# Import Libraries

```{r}
library(rethinking)
library(tidyverse)
```

# 4.1 - Why normal dists are normal

```{r}
#Simulate flipping a coin 16 times and walking depending on the outcome

pos <- replicate(1000, sum(runif(16,-1,1)))

hist(pos)
```

## 4.1.2 Normality by multiplication

```{r}
growth <- replicate(10000, prod(1+runif(12,0,0.1)))

dens(growth, norm.comp=TRUE)
```

## 4.1.3 Normal by log-multiplication

```{r}
log.big <- replicate(10000, log(prod(1+runif(12,0,0.5))))
dens(log.big)
```

# 4.2 - A language for describing models

1.  We recognize a set of variables to work with. Some of these are observable, and are called *data.* Others are unobservable things like rates and averages, which are called *parameters*.
2.  We define each variable either in terms of the other variables or in terms of a probability distribution.
3.  The combination of variables and their probability distributions defines a *joint generative model* that can be used both to simulate hypothetical observations as well as analyze real ones.

W \~ Binomial(*N,p*) - Read as, "The count *W* is distributed binomially with sample size *N* and probability *p*.

p \~ Uniform(0,1) - Read as, "The prior for *p* is assumed to be uniform between zero and one.

Both of the above lines in the model are stochastic, as indicated by the \~. A stochastic relationship is just a mapping of a variable or parameter onto a distribution. It is stochastic because no single instance of he variable on the left is known with certainty.

# 4.3 - Gaussian model of height

```{r}
data(Howell1)
d <- Howell1
str(d)
#height - cm's
#weight - kg's
#age - years
#gender - 1=Male, 0=Female
```

```{r}
precis(d)
```

```{r}
#Book only wants to work with adults
d2 <- d[d$age > 18,]
```

## 4.3.2 The model

```{r}
dens(d2$height)
```

Height is approximately normal so we can describe it as: $h_i \sim$ Normal($\mu,\sigma$).

We need priors which the book gives us as: $\mu\sim Normal(178,20)$ and $\sigma\sim Uniform(0,50)$.

```{r}
#Plot the priors
curve(dnorm(x, 178, 20), from=100, to=250)
```

```{r}
curve(dunif(x,0,50), from=-10, to=60)
```

```{r}
#Prior Predictive simulation
sample_mu <- rnorm(1e4, 178, 20)
sample_sigma <- runif(1e4, 0, 50)
prior_h <- rnorm(1e4, sample_mu, sample_sigma)
dens(prior_h)
```

## 4.3.3 Grid approximation of the posterior

```{r}
#Grid approximation of the posterior
mu.list <- seq(from=150, to=160, length.out=100)
sigma.list <- seq(from=7, to=9, length.out=100)
post <- expand.grid(mu=mu.list, sigma=sigma.list)
post$LL <- sapply(1:nrow(post), function(i) sum(
  dnorm(d2$height, post$mu[i], post$sigma[i], log=TRUE)))
post$prod <- post$LL + dnorm(post$mu, 178, 20, TRUE) + dunif(post$sigma, 0, 50, TRUE)
post$prob <- exp(post$prod - max(post$prod))
image_xyz(post$mu, post$sigma, post$prob)
```

## 4.3.4 Sampling from the posterior

```{r}
sample.rows <- sample(1:nrow(post), size=1e4, replace=TRUE, prob=post$prob)
sample.mu <- post$mu[sample.rows]
sample.sigma <- post$sigma[sample.rows]
plot(sample.mu, sample.sigma, cex=0.5, pch=16, col=col.alpha(rangi2,0.1))
```

```{r}
#Marginal posterior density of mu
dens(sample.mu)
```

```{r}
#Marginal posterior density of sigma
dens(sample.sigma)
```

```{r}
PI(sample.mu)
PI(sample.sigma)
```

## 4.3.5 Finding the posterior distribution with `quap`

```{r}
flist <- alist(
  height ~ dnorm(mu, sigma),
  mu ~ dnorm(178, 20),
  sigma ~ dunif(0,50)
)

m4.1 <- quap(flist, data=d2)

precis(m4.1)
```

These numbers provide Gaussian approximations for each parameter's marginal distributions. This means the plausibility of each value of $\mu$ after averaging over the plausibilities of each value of $\sigma$, is given by a Gaussian distribution with mean 154.6 and standard deviation of 0.4.

We can change our prior of $\mu$ to be much more concentrated around our mean, signalling that it is much stronger.

```{r}
m4.2 <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu ~ dnorm(178, 0.1),
    sigma ~ dunif(0,50)
  ), data = d2
)

precis(m4.2)
```

Our estimate for $\mu$ has hardly changed off the prior now because the prior was so much more concentrated. However, our estimate for $\sigma$ has changed a lot even though we didn't change its prior. This is because once the model (a metaphorical golem) is certain the mean is near 178 - as the prior insists - then the model has to estimate a $\sigma$ conditional on that fact. Which results in a different posterior for $\sigma$.

## 4.3.6 Sampling from `quap`

```{r}
diag(vcov(m4.1))
cov2cor(vcov(m4.1))
```

The two element vector (listed first) is the list of variances. Taking the square root of the vector will give the standard deviations shown in the `precis` output of `m4.1`. While the two-by-two matrix represents the correlation (bounded by +1 and -1) between the pairs of parameters. The estimates near 0 for both tell us that learning $\mu$ tells us nothing about $\sigma$ and vice versa.

```{r}
#Sampling from multidimensional posterior
post <- extract.samples(m4.1, n=1e4)
precis(post)
```

# 4.4 - Linear prediction

```{r}
plot(d2$height, d2$weight)
```

## 4.4.1.3 *Priors*

```{r}
set.seed(2971)
N <- 100
a <- rnorm(N, 178, 20)
b <- rnorm(N, 0, 10)
```

```{r}
plot(NULL, xlim=range(d2$weight), ylim=c(-100,400), xlab="weight", ylab="height")
abline(h=0, lty=2)
abline(h=272, lty=1, lwd=0.5)
mtext("b ~ dnorm(0,10)")
xbar <- mean(d2$weight)
for(i in 1:N) curve(a[i] + b[i]*(x-xbar), from=min(d2$weight), to=max(d2$weight), add=TRUE, col=col.alpha("black", 0.2))
```

```{r}
#Experimenting with b being log-normal
b <- rlnorm(1e4, 0, 1)
dens(b, xlim=c(0,5), adj=0.1)
```

```{r}
set.seed(2971)
N <- 100
a <- rnorm(N, 178, 20)
b <- rlnorm(N, 0, 1)

plot(NULL, xlim=range(d2$weight), ylim=c(-100,400), xlab="weight", ylab="height")
abline(h=0, lty=2)
abline(h=272, lty=1, lwd=0.5)
mtext("log(b) ~ dnorm(0,1)")
xbar <- mean(d2$weight)
for(i in 1:N) curve(a[i] + b[i]*(x-xbar), from=min(d2$weight), to=max(d2$weight), add=TRUE, col=col.alpha("black", 0.2))
```

## 4.4.2 Finding the posterior distribution

```{r}
#Load data again
data("Howell1"); d <- Howell1; d2 <- d[d$age > 18,]

#defin average weight
xbar <- mean(d2$weight)

#fit model
m4.3 <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b*(weight - xbar),
    a ~ dnorm(178,20),
    b ~ dlnorm(0,1),
    sigma ~ dunif(0,50)
  ), data=d2
)
```

### 4.4.3.1 Tables of marginal distributions

```{r}
precis(m4.3)
```

The first row gives us the quadratic approximation for $\alpha$, the second row $\beta$, and the third row $\sigma$.

```{r}
round(vcov(m4.3),3)
```

### 4.4.3.2 Plotting posterior inference against the data

```{r}
plot(height ~ weight, data=d2, col=rangi2)
post <- extract.samples(m4.3)
a_map <- mean(post$a)
b_map <- mean(post$b)
curve(a_map + b_map*(x-xbar), add=TRUE)
```

### 4.4.3.4 Plotting regression intervals and contours

General recipe for generating predictions and intervals from the posterior of a fit model:

1.  Use `link` to generate distributions of posterior values for $\mu$. The default behavior of `link` is to use the original data, so you have to pass it a list of new horizontal axis values you want to plot posterior predictions across.
2.  Use summary functions like `mean` or `PI` to find averages and lower and upper bounds of $\mu$ for each value of the predictor variable
3.  Finally, use plotting functions like `lines` and shade to draw the lines and intervals. Or plot the distributions of predictions, or do further numerical calculations with them.

# 4.5 Curves from lines

## 4.5.1 Polynomial regression

```{r}
plot(height ~ weight, d)
```

```{r}
#Very different than in the book
data("Howell1")
d <- Howell1

d$weight_s <- (d$weight - mean(d$weight))/sd(d$weight)
d$weight_s2 <- d$weight^2

m4.5 <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b1*weight_s + b2*weight_s2,
    a ~ dnorm(178, 20),
    b1 <- dlnorm(0,1),
    b2 ~ dnorm(0,1),
    sigma ~ dunif(0,50)
  ), data=d
)
precis(m4.5)
```

```{r}
weight.seq <- seq(from=-2.2, to=2, length.out=30)
pred_dat <- list(weight_s=weight.seq, weight_s2=weight.seq^2)
mu <- link(m4.5, data=pred_dat)
mu.mean <- apply(mu, 2, mean)
mu.PI <- apply(mu, 2, PI, prob=0.89)
sim.height <- sim(m4.5, data=pred_dat)
height.PI <- apply(sim.height, 2, PI, prob=0.89)
```

```{r}
#Where is the line?
plot(height ~ weight, d, col=col.alpha(rangi2, 0.5))
lines(weight.seq, mu.mean)
shade(mu.PI, weight.seq)
shade(height.PI, weight.seq)
```