### 2023-09-17 Stats Rethinking/Ch4/Dr_Parson_ch_4_problems.qmd
Finish Chapter 4 problems
I finished the chapter 4 problem, including all the splines questions.

---
title: "Chapter 4 practice"
format: html
editor: visual
---

```{r}
library(tidyverse)
```

# E problems

## 4E1

The likelihood is y_i \~ Normal(mu, sigma)

## 4E2

There are two parameters in the posterior distribution, mu and sigma.

## 4E3

Bayes' theorem gives us the following formula for the posterior joint density of $\mu$ and $\sigma$.

$$
Pr(\mu, \sigma | y_1, \dots, y_n) = \frac{\prod_{i = 1}^n Pr(y_i | \mu, \sigma) \operatorname{Normal}(\mu | 0, 10) \operatorname{Exponential}(\sigma | 1)}
{\int\int \prod_{i = 1}^n Pr(y_i | \mu, \sigma) \operatorname{Normal}(\mu | 0, 10)\operatorname{Exponential}(\sigma | 1) d\mu d\sigma}
$$

## 4E4

The linear model is $\mu_i = \alpha + \beta x_i$.

## 4E5

There are three parameters in the posterior distribution.

# M problems

## 4M1

We have the following model:

y_i \~ Normal(mu, sigma)

mu \~ Normal(0, 10)

sigma \~ Exponential(1)

To simulate data from the prior distribution, we first draw 10,000 samples (say) for mu and sigma from the given prior distributions and then draw 10,000 samples (using our mu and sigma samples) from the normal distribution.

```{r}
N <- 1e4
mus <- rnorm(N, mean = 0, sd = 10)
sigmas <- rexp(N, rate = 1)
ys <- rnorm(N, mean = mus, sd = sigmas)
```

We can make a density graph of the simulated y values. For comparison, I have overlaid a normal distribution with the same mean and standard deviation as the list of simulated y values. It looks pretty close! (Checking whether the prior distribution of the y values is actually a normal distribution requires some probability theory. Maybe something to do at the end of MATH 351...)

```{r}
ggplot() +
  geom_density(mapping = aes(x = ys)) +
  geom_function(fun = dnorm, args = list(mean = mean(ys), sd = sd(ys)), color = "red")
```

## 4M2

In `quap` notation, the model looks like this:

```         
y ~ dnorm(mu, sigma),
mu ~ dnorm(0, 10),
sigma ~ dexp(1)
```

## 4M3

The following mathematical model definition matches the given `quap` formula:

y_i \~ Normal(mu, sigma)

mu_i = a + b x_i

a \~ Normal(0, 10)

b \~ Uniform(0, 1)

sigma \~ Exp(1)

## 4M4

| Student | year | height |
|---------|------|--------|
| 1       | 0    | 60     |
| 1       | 1    | 61     |
| 1       | 2    | 62     |
| 2       | 0    | 65     |
| 2       | 1    | 67     |
| 2       | 2    | 69     |
| 3       | 0    | 55     |
| 3       | 1    | 60     |
| 3       | 2    | 65     |

In our model, we ignore the "student" column. We model the height column using the year column as predictor. The subscript i indicates the row in the data table. The parameter a is something like the average height in year 0. The parameter b is something like the typical increase in height per year.

height_i \~ Normal(mu_i, sigma)

m_i = a + b year_i

sigma \~ Exp(2)

a \~ Normal(5, 0.5)

b \~ Exp(6)

Likelihood: heights for the students could be normally distributed or a mixture of normal distributions.

Prior on sigma: I am imagining students in middle school (6th, 7th, 8th). Typical differences in height might be 6 inches (1/2 foot). We'll use an exponential distribution.

Prior on a: This would be something like average height among 6th graders. Maybe 5 feet give or take 6 inches to be safe.

Prior on b: This would be something like average height increase per year. This parameter could not be negative! Maybe 2 inches (1/6 foot) would be a reasonable guess for middle school. We'll use an exponential distribution to guarantee positive values. But we could try a log-normal distribution instead.

A richer (multilevel) model would allow for different slopes for each student with pooling. Something like this might work. (But I haven't thought about this at all!)

height_ij \~ Normal(mu_ij, sigma)

mu_ij = a_i + b_i year_j

sigma \~ Exp(2)

a_i \~ Normal(5, 0.5)

b_i \~ Exp(6)

```{r}
N <- 1e1
sigmas <- rexp(N, 2)
as <- rnorm(N, 5, 0.5)
bs <- rexp(N, 6)

student_simulation <- tibble(replication = rep(1:N, times = 3),
                             year = rep(c(0,1,2), each = N),
                             a = rep(as, times = 3),
                             b = rep(bs, times = 3),
                             sigma = rep(sigmas, times = 3),
                             height = rnorm(3*N, mean = a + b*year, sd = sigma))
```

```{r}
ggplot(data = student_simulation) +
  geom_point(mapping = aes(x = factor(year), y = height)) +
  geom_line(mapping = aes(x = factor(year), y = height, group = replication))
```

Huh! Sometimes students shrink, according to this model. To make this work better, one might consider a time-series model instead of a generic regression model with time as a predictor.

## 4M5

The model above already includes a prior for the slope that does not allow negative slopes. The problem with the model is that it does not remember how tall students are from year to year. The random component of the prediction (likelihood) is intended to reflect between-student variability in heights. But that leads to occasional shrinking when we track one person.

I guess that's not how we should use the model. We cannot model the growth of a singular person with this set-up. What we are modeling is the heights of sixth graders, seventh graders, and eighth graders as groups.

The following chunks simulate 100 sixth graders, 100 seventh graders, and 100 eighth graders. We can see that the prior is not great, since it has too much variability.

```{r}
N <- 1e2
sigmas <- rexp(3*N, 2)
as <- rnorm(3*N, 5, 0.5)
bs <- rexp(3*N, 6)

student_simulation2 <- tibble(year = rep(c(0,1,2), each = N),
                             a = as,
                             b = bs,
                             sigma = sigmas,
                             height = rnorm(3*N, mean = a + b*year, sd = sigma))
```

```{r}
ggplot(data = student_simulation2) +
  geom_jitter(mapping = aes(x = factor(year), y = height), height = 0, width = 0.1, alpha = 0.2)
```

The confusing part of the problem statement is the tracking of heights over three years. It might be clearer framed as data about students in a middle school.

## 4M6

The word "variance" is a strange choice! I think what McElreath has in mind, though, is that we could limit sigma above in some way by 64 cm. I'm not sure exactly what "variance" means here. Maybe we could use sigma \~ Uniform(0, 64 / (2.54 \* 12)).

## 4M7

Refitting model changes `a`, as one would expect, but the model predictions are the same.

*Regression and Other Stories* (<https://avehtari.github.io/ROS-Examples/>) suggests centering predictors so that one can assign meaningful priors.

```{r}
library(rethinking)
data(Howell1)
```

```{r}
d2 <- Howell1 %>% filter(age >= 18)

xbar <- d2 %>% summarize(m = mean(weight)) %>% pull(m)
```

```{r}
m4.3 <- quap(
  alist(
    height ~ dnorm( mu , sigma ) ,
    mu <- a + b*( weight - xbar ) ,
    a ~ dnorm( 178 , 20 ) ,
    b ~ dlnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 50 )
    ) , data=d2 )

precis(m4.3)
```

```{r}
post <- extract.samples(m4.3)

post %>% pull(a) %>% quantile(c(0.05, 0.95))
```

```{r}
PI(post %>% pull(a), prob = 0.9)
```

```{r}
m4.3v2 <- quap(
  alist(
    height ~ dnorm( mu , sigma ) ,
    mu <- a + b * weight ,
    a ~ dnorm( 178 , 20 ) ,
    b ~ dnorm( 1 , 2 ) ,
    sigma ~ dunif( 0 , 50 )
    ) , data=d2 )
```

```{r}
post2 <- extract.samples(m4.3v2)
vcov(m4.3v2)
```

```{r}
sim.height <- sim(m4.3, data = list(weight = 50))
ggplot() + 
  geom_density(mapping = aes(x = sim.height))
```

```{r}
quantile(sim.height, c(0.025, 0.975))
```

```{r}
sim.heightv2 <- sim(m4.3v2, data = list(weight = 50))
ggplot() + 
  geom_density(mapping = aes(x = sim.heightv2))
```

```{r}
quantile(sim.heightv2, c(0.025, 0.975))
```

```{r}
m4.3v3 <- quap(
  alist(
    height ~ dnorm( mu , sigma ) ,
    mu <- a + b*( weight - xbar ) + c * male ,
    mu2 <- a + b*(50 - xbar),
    a ~ dnorm( 178 , 20 ) ,
    b ~ dlnorm( 0 , 1 ) ,
    c ~ dnorm( 0, 1 ),
    sigma ~ dunif( 0 , 50 )
    ) , data=d2 )

post <- extract.samples(m4.3v3)

post %>% pull(c) %>% quantile(c(0.05, 0.95))
```

## 4M8

I'll have to come back to this one later! I am feeling too lazy to study the syntax for fitting spline models at the moment.

For general information on splines (and polynomial fits) see Chapter 7 of *Statistical Learning* (<https://www.statlearning.com>).

```{r}
data("cherry_blossoms")

ggplot(data = cherry_blossoms) +
  geom_point(mapping = aes(x = year, y = doy))
```

```{r}
cherry_full <- cherry_blossoms %>% drop_na(doy)
num_knots <- 15
knot_list <- quantile( cherry_full %>% pull(year) , probs=seq(0, 1, length.out=num_knots) )
```

```{r}
library(splines)
B <- bs(cherry_full %>% pull(year),
        knots = knot_list[-c(1,num_knots)], degree=3 , intercept=TRUE )
cherry_full <- cherry_blossoms %>% drop_na(doy)
num_knots <- 15
knot_list <- quantile( cherry_full %>% pull(year) , probs=seq(0, 1, length.out=num_knots) )
```

The following chunk is copied from the textbook.

```{r}
plot( NULL , xlim=range(cherry_full$year) , ylim=c(0,1) , xlab="year" , ylab="basis" )
for ( i in 1:ncol(B) ) lines( cherry_full$year , B[,i] )
```

```{r}
m4.7 <- quap(
  alist(
    D ~ dnorm( mu , sigma ) ,
    mu <- a + B %*% w ,
    a ~ dnorm(100,10),
    w ~ dnorm(0,100),
    sigma ~ dexp(1)
    ),
  data=list( D= cherry_full %>% pull(doy), B = B ) ,
  start=list( w=rep( 0 , ncol(B) ) ) )
```

```{r}
post <- extract.samples( m4.7 )
w <- apply( post$w , 2 , mean )
plot( NULL , xlim = range(cherry_full$year) ,
      ylim=c(-6,6) , xlab="year" , ylab="basis * weight" )
for ( i in 1:ncol(B) ) lines( cherry_full$year , w[i]*B[,i] )
```

```{r}
mu <- link( m4.7 )
mu_PI <- apply(mu,2,PI,0.97)
plot( cherry_full$year , cherry_full$doy , col=col.alpha(rangi2, 0.3) , pch=16, 
      shade( mu_PI , cherry_full$year , col=col.alpha("black",0.5) ))
```

With 100 knots, the model is very flexible and matches data fluctuations on small scales.

With 5 knots, the model is not flexible at all and only reflects long-term trends. (I like 5 knots!)

Increasing the standard deviation on the prior for the weights (say from 10 to 100) also increases the flexibility of the model but not as dramatically as increasing the number of knots.

# H problems

## 4H1

There should be a better way to organize displaying this work! If only the commands were made to give tidy outputs...

```{r}
newdata <- c(46.95, 43.72, 64.78, 32.59, 54.63)
sim.height.new <- sim(m4.3v2, data = list(weight = newdata))

for(i in 1:5){
  print(round(mean(sim.height.new[, i]), 1))
  print(round(PI(sim.height.new[,i]), 1))
}
```

This question is somewhat silly. Who measures the weight of an adult to within 10 grams? Why bother when the resulting prediction intervals for height are so wide?

## 4H2

### Part a

We should do all the manipulations using `tidyverse`! I recommend barring `[ ] $`.

```{r}
howell_young <- Howell1 %>% filter(age < 18)
```

```{r}
young_fit <- quap(
  alist(height ~ dnorm(mu, sigma),
        mu <- a + b * (weight - mean(weight)),
        a ~ dnorm(100, 20),
        b ~ dlnorm(1, 1),
        sigma ~ dunif(0, 50)
        ), data = howell_young
)
```

```{r}
precis(young_fit)
```

For comparison, we can fit a traditional `lm` model. The coefficients and standard errors are essentially the same! Since we have so much data, the priors do not much matter. (And the `quap` approximation also works like `lm`.)

```{r}
howell_young <- howell_young %>% mutate(c.weight = weight - mean(weight))
lm(height ~ c.weight, data = howell_young) %>% summary
```

A 10 kg difference in weight is associated with a 27 cm difference in height.

### Part b

```{r}
xbar <- howell_young %>% summarize(m = mean(weight)) %>% pull(m)
```

```{r}
samples <- extract.samples(young_fit)
N <- 20
xs <- seq(from = 4, to = 45, length.out = N) - xbar
mu_min <- rep(0, times = N)
mu_max <- rep(0, times = N)
pred_min <- rep(0, times = N)
pred_max <- rep(0, times = N)

for (i in 1:20){
  q <- samples %>% mutate(mu = a + b * xs[i],
                     pred = rnorm(nrow(samples), mu, sigma)) %>%
    summarize(q_mu_low = quantile(mu, 0.055),
              q_mu_high = quantile(mu, 0.945),
              q_pred_low = quantile(pred, 0.055),
              q_pred_high = quantile(pred, 0.945))
  mu_min[i] <- q %>% pull(q_mu_low)
  mu_max[i] <- q %>% pull(q_mu_high)
  pred_min[i] <- q %>% pull(q_pred_low)
  pred_max[i] <- q %>% pull(q_pred_high)
}

strips <- tibble(mu_min = mu_min, mu_max = mu_max,
                 pred_min = pred_min, pred_max = pred_max)
```

```{r}
ggplot(data = howell_young) +
  geom_ribbon(data = strips,
              mapping = aes(x = xs + xbar, ymin = mu_min, ymax = mu_max),
              alpha = 0.2) +
  geom_ribbon(data = strips,
              mapping = aes(x = xs + xbar, ymin = pred_min, ymax = pred_max),
              alpha = 0.3) +
  geom_point(mapping = aes(x = weight, y = height), shape = 1, color = "blue", alpha = 0.5) +
  geom_function(fun = function(x){108.31 + 2.72 * (x - xbar)}) +
  labs(x = "weight (kg)", y = "height (cm)") +
  theme_bw()
```

This model does not fit the data well. The scatterplot shows clear non-linearity. Many observations at the low and high ends of weight fall outside of the prediction interval. The conditional means clearly do not fall in the (very narrow) interval for the means.

### Part c

We need to use a non-linear model!

## 4H3

I did not know what to do for the prior on this one. My first choice was horrible, and the posterior model did not fit the data well. I checked the model from `lm` (matching a "non-informative prior") and adjusted my prior to be much more flexible. This model now essentially reproduces `lm`.

```{r}
log_fit <- quap(
  alist(height ~ dnorm(mu, sigma),
        mu <- a + b * log(weight),
        a ~ dnorm(0, 20),
        b ~ dnorm(0, 40),
        sigma ~ dunif(0, 50)),
  data = Howell1
)

precis(log_fit)
```

```{r}
ggplot(data = Howell1) +
  geom_point(mapping = aes(x = weight, y = height)) +
  geom_function(fun = function(x)(-23.70 + 47.05 * log(x)))
```

```{r}
lm(height ~ log(weight), data = Howell1) %>% summary
```

The following code produces the 97% interval for the mean and the 97% interval for the predicted heights.

```{r}
samples <- extract.samples(log_fit)
N <- 20
xs <- seq(from = 4, to = 63, length.out = N)
mu_min <- rep(0, times = N)
mu_max <- rep(0, times = N)
pred_min <- rep(0, times = N)
pred_max <- rep(0, times = N)

for (i in 1:20){
  q <- samples %>% mutate(mu = a + b * log(xs[i]),
                     pred = rnorm(nrow(samples), mu, sigma)) %>%
    summarize(q_mu_low = quantile(mu, 0.015),
              q_mu_high = quantile(mu, 0.985),
              q_pred_low = quantile(pred, 0.015),
              q_pred_high = quantile(pred, 0.985))
  mu_min[i] <- q %>% pull(q_mu_low)
  mu_max[i] <- q %>% pull(q_mu_high)
  pred_min[i] <- q %>% pull(q_pred_low)
  pred_max[i] <- q %>% pull(q_pred_high)
}

strips <- tibble(mu_min = mu_min, mu_max = mu_max,
                 pred_min = pred_min, pred_max = pred_max)
```

```{r}
ggplot(data = Howell1) +
  geom_ribbon(data = strips,
              mapping = aes(x = xs, ymin = mu_min, ymax = mu_max),
              alpha = 0.4) +
  geom_ribbon(data = strips,
              mapping = aes(x = xs, ymin = pred_min, ymax = pred_max),
              alpha = 0.3) +
  geom_point(mapping = aes(x = weight, y = height), shape = 1, color = "blue", alpha = 0.5) +
  geom_function(fun = function(x)(-23.70 + 47.05 * log(x))) +
  labs(x = "weight (kg)", y = "height (cm)") +
  theme_bw()
```

This model looks much better, though the prediction intervals seem too wide for low weights and to narrow for high weights. Perhaps `sigma` could vary with the weight...

## 4H4

The quadratic model has the following likelihood:

height \~ Normal(mu, sigma)

mu = a + b_1 weight_s + b_2 weight_s\^2,

where weight_s is standardized (z-score) weight.

The following chunk simulates data for various choices of prior.

```{r}
N = nrow(Howell1)
howell_quadratic <- Howell1 %>% mutate(weight_s = (weight - mean(weight)) / sd(weight),
                                       weight_s2 = weight_s^2,
                                       a = rnorm(N, 178, 20),
                                       b1 = rlnorm(N, 2.5, 1),
                                       b2 = rnorm(N, -20, 1),
                                       sigma = runif(N, 0, 10),
                                       mu = a + b1 * weight_s + b2 * weight_s^2,
                                       height_sim = rnorm(N, mu, sigma)
                                       )
```

```{r}
ggplot(data = howell_quadratic) +
  geom_point(mapping = aes(x = weight, y = height_sim))
```

One could keep on fiddling... but the text asks us not to fit a model by hand.

## 4H5

```{r}
cherry_temp <- cherry_blossoms %>% drop_na(doy, temp)

ggplot(data = cherry_temp) +
  geom_point(mapping = aes(x = temp, y = doy))
```

```{r}
linear_fit <- quap(
  alist(doy ~ dnorm(mu, sigma),
        mu <- a + b * (temp - mean(temp)),
        a ~ dnorm(100, 20),
        b ~ dnorm(10, 10),
        sigma ~ dexp(1/10)),
  data = cherry_temp
)

precis(linear_fit)
```

Perhaps I should have started with a prior predictive check.

```{r}
N <- nrow(cherry_temp)
prior_predictive <- cherry_temp %>% mutate(a = rnorm(N, 100, 20),
                                           b = rnorm(N, 10, 10),
                                           sigma = rexp(N, 1/10),
                                           mu = a + b * (temp - mean(temp)),
                                           sim_day = rnorm(N, mu, sigma))

ggplot(data = prior_predictive) +
  geom_point(mapping = aes(x = temp, y = sim_day))
```

This looks fine, I think, for a prior model. There is too much variability---sometimes we predict blossom before day 0!\--but it looks like a reasonable starting point.

We can try a posterior predictive check as well.

```{r}
post <- extract.samples(linear_fit)

posterior_predictive <- cherry_temp %>% cbind(slice_head(post, n = N)) %>%
  mutate(mu = a + b * (temp - mean(temp)),
         doy_sim = rnorm(N, mu, sigma))

ggplot(data = posterior_predictive) +
  geom_point(mapping = aes(x = temp, y = doy_sim))
```

```{r}
samples <- extract.samples(linear_fit)
N <- 20
xs <- seq(from = 4.5, to = 8.5, length.out = N)
mu_min <- rep(0, times = N)
mu_max <- rep(0, times = N)
pred_min <- rep(0, times = N)
pred_max <- rep(0, times = N)
xbar <- cherry_temp %>% summarize(m = mean(temp)) %>% pull(m)

for (i in 1:20){
  q <- samples %>% mutate(mu = a + b * (xs[i] - xbar),
                     pred = rnorm(nrow(samples), mu, sigma)) %>%
    summarize(q_mu_low = quantile(mu, 0.015),
              q_mu_high = quantile(mu, 0.985),
              q_pred_low = quantile(pred, 0.015),
              q_pred_high = quantile(pred, 0.985))
  mu_min[i] <- q %>% pull(q_mu_low)
  mu_max[i] <- q %>% pull(q_mu_high)
  pred_min[i] <- q %>% pull(q_pred_low)
  pred_max[i] <- q %>% pull(q_pred_high)
}

strips <- tibble(mu_min = mu_min, mu_max = mu_max,
                 pred_min = pred_min, pred_max = pred_max)
```

```{r}
ggplot(data = cherry_temp) +
  geom_ribbon(data = strips,
              mapping = aes(x = xs, ymin = mu_min, ymax = mu_max),
              alpha = 0.4) +
  geom_ribbon(data = strips,
              mapping = aes(x = xs, ymin = pred_min, ymax = pred_max),
              alpha = 0.3) +
  geom_point(mapping = aes(x = temp, y = doy), shape = 1, color = "blue", alpha = 0.5) +
  geom_function(fun = function(x)(104.92 + (-2.98) * (x - xbar))) +
  labs(x = "March temperature (C)",
       y = "Blossom date (day of year") +
  theme_bw()
```

I am content with this linear model!

## 4H6

```{r}
cherry_full <- cherry_blossoms %>% drop_na(doy)
num_knots <- 15
knot_list <- quantile( cherry_full %>% pull(year) , probs=seq(0, 1, length.out=num_knots) )
```

```{r}
library(splines)
B <- bs(cherry_full %>% pull(year),
        knots = knot_list[-c(1,num_knots)], degree=3 , intercept=TRUE )
cherry_full <- cherry_blossoms %>% drop_na(doy)
```

```{r}
cherry_refit <- quap(
  alist(
    D ~ dnorm( mu , sigma ) ,
    mu <- a + B %*% w ,
    a ~ dnorm(100,10),
    w ~ dnorm(0,10),
    sigma ~ dexp(1)
    ),
  data=list( D= cherry_full %>% pull(doy), B = B ) ,
  start=list( w=rep( 0 , ncol(B) ) ) )

precis(cherry_refit)
```

```{r}
post <- extract.samples( cherry_refit )
w <- apply( post$w , 2 , mean )
plot( NULL , xlim = range(cherry_full$year) ,
      ylim=c(-6,6) , xlab="year" , ylab="basis * weight" )
for ( i in 1:ncol(B) ) lines( cherry_full$year , w[i]*B[,i] )
```

```{r}
mu <- link( cherry_refit )
mu_PI <- apply(mu,2,PI,0.97)
plot( cherry_full$year , cherry_full$doy , col=col.alpha(rangi2, 0.3) , pch=16, 
      shade( mu_PI , cherry_full$year , col=col.alpha("black",0.5) ))
```

Not sure what this question is getting at. When I change the mean on the prior for the weights, the fitting procedure compensates by adjusting a. I end up with essentially the same final model.

## 4H7

What happened to 4H7?

## 4H8

We refit the model without an intercept term. To do that, we change the call to "bs", using `intercept = FALSE` and including all the knots (instead of dropping the first).

```{r}
cherry_full <- cherry_blossoms %>% drop_na(doy)
num_knots <- 15
knot_list <- quantile( cherry_full %>% pull(year) , probs=seq(0, 1, length.out=num_knots) )
```

```{r}
library(splines)
B <- bs(cherry_full %>% pull(year),
        knots = knot_list, degree=3 , intercept = FALSE )
cherry_full <- cherry_blossoms %>% drop_na(doy)
```

Removing the intercept means that we need to change the priors. I'm not sure what to expect, but we can set the means for the spline terms to the old mean for the intercept.

```{r}
cherry_no_intercept <- quap(
  alist(
    D ~ dnorm( mu , sigma ) ,
    mu <- B %*% w ,
    w ~ dnorm(100,100),
    sigma ~ dexp(1)
    ),
  data=list( D= cherry_full %>% pull(doy), B = B ) ,
  start=list( w=rep( 100 , ncol(B) ) ) )

precis(cherry_no_intercept, depth = 2)
```

Here are the spline terms, with large weights, as expected.

```{r}
post <- extract.samples( cherry_no_intercept )
w <- apply( post$w , 2 , mean )
plot( NULL , xlim = range(cherry_full$year) ,
      ylim=c(0,120) , xlab="year" , ylab="basis * weight" )
for ( i in 1:ncol(B) ) lines( cherry_full$year , w[i]*B[,i] )
```

And here is the model, the same as before.

```{r}
mu <- link( cherry_no_intercept )
mu_PI <- apply(mu,2,PI,0.97)
plot( cherry_full$year , cherry_full$doy , col=col.alpha(rangi2, 0.3) , pch=16, 
      shade( mu_PI , cherry_full$year , col=col.alpha("black",0.5) ))
```