### 2023-09-21 Stats Rethinking/Ch7/Practice Problems.qmd
Created Practice Problems Sheet

---
title: "Practice Problems"
format: html
---

# Load Libraries

```{r}
library(rethinking)
library(tidyverse)
```

# 7E1