### 2023-10-03 Stats Rethinking/Ch9/Practice Problems.qmd
Created chpt 9 practice doc

---
title: "Practice Problems"
format: html
---

# Load Libraries

```{r}
library(tidyverse)
library(rethinking)
```

# 9E1