### 2023-10-04 Lahman Analysis/Rough Analysis/RoughAnalysis.qmd
Got stan set up

---
title: "Rough Analysis"
format: html
---

# Load Libraries

## Installing Lahman

```{r}
#| eval: false

#Installing Lahman is easy and simply requires running this code chunk

install.packages("Lahman")
library(Lahman)
```

## Loading Lahman

```{r}
#There are more datasets in Lahman than we care about so we'll select "Batting"

Batting <- Lahman::Batting
```

The `Batting` data set describes hitting outcomes for each MLB player each year from 1871-present. The explanation of the variable abbreviations is available on page 14 of [this document](https://cran.r-project.org/web/packages/Lahman/Lahman.pdf).

Of course, we don't want to analyze all data from 1871-present since the game has changed considerable since then. So filter to only consider seasons past WW2 (1945). We can always cut down from here, but this will give us a huge sample to use where the rules (nor the game) has changed significantly since.

```{r}
Modern.Batting <- Batting %>% 
  filter(yearID > 1945)
```

## Load Other Libraries

```{r}
#Of course we need the other two main libraries
library(tidyverse)
library(rethinking)
```

# Quap

# Ulam