### 2023-10-11 Stats Rethinking/Ch11/Practice Problems.qmd
Added space for chpt 11 practice problems

---
title: "Practice Problems"
format: html
---

# Load Libraries

```{r}
library(tidyverse)
library(rethinking)
```

# 11E1

# 11E2

# 11E3

# 11E4

# 11M1

# 11M2

# 11M3

# 11M4

# 11H1

# 11H2

# 11H3

# 11H4