### 2023-10-12 Lahman Analysis/Rough Analysis/BinomialRoughAnalysis.qmd
Added space for Binomial regression analysis

---
title: "Binomial Rough Analysis"
format: html
---

# Load Libraries

```{r}
library(tidyverse)
library(rethinking)
library(Lahman)
```

# Clean Lahman

```{r}
Batting <- Lahman::Batting

Batting %>% 
  filter(yearID >= 1945)
```

# Quap

# Ulam

# Comparison