### 2023-10-12 Lahman Analysis/Rough Analysis/BinomialRoughAnalysis.qmd
Finished through medium practice problems

---
title: "Binomial Rough Analysis"
format: html
---

# Load Libraries

```{r}
library(tidyverse)
library(rethinking)
library(Lahman)
```

# Clean Lahman

```{r}
Batting <- Lahman::Batting

Batting <- Batting %>% 
  filter(yearID >= 1945) %>% 
  select(playerID, yearID, lgID, AB, AB, HR) %>% 
  mutate(lg = ifelse(lgID=="NL",1,2))
```

# Quap

# Ulam

```{r}
head(Batting)
```

```{r}
slim <- list(
  AB = Batting$AB,
  HR = Batting$HR,
  LG = Batting$lg
)

str(slim)
```

```{r}
u1 <- ulam(
  alist(
    HR ~ binomial(p),
    logit(p) = ,
  ), data=slim, chains=4, cores=4
)
```

# Comparison