### 2023-10-19 Lahman Analysis/Rough Analysis/CleanLahman.qmd
Started more complex model

---
title: "Clean Lahman"
format: html
---

# Load Libraries

```{r}
#| message: false

library(tidyverse)
library(Lahman)
```

# Clean

```{r}
#| message: false

#Creates a DF called "Batting" which has players stats from 1973-2019

#Load the DF's from Lahman directly
Batting <- Lahman::Batting
Teams <- Lahman::Teams
People <- Lahman::People

Batting <- Batting %>% 
  #Minimum year = DH was introduced; maximum year = before COVID
  filter(yearID >= 1973 & yearID <= 2019)

#Filter teams for years considered and select only necessary columns
Teams <- Teams %>% 
  filter(yearID>=1973 & yearID <= 2019) %>% 
  select(yearID, teamID, park)

#Combine current "Batting" and "Teams" DF to associate player with stadium played at
Batting <- Batting %>% 
  right_join(Teams, by = join_by(yearID, teamID))

#Add player's age to the "Batting" DF
Battin <- Batting %>% 
  right_join(Lahman::People, by = join_by(playerID)) %>% 
  mutate(Age = yearID - birthYear, .after=yearID) %>%
  #Don't consider the people who don't have a birthday in the database
  filter(!is.na(Age))
```

# Cleaning

```{r}
Batting <- Lahman::Batting

Batting <- Batting %>% 
  #Minimum year = DH was introduced; maximum year = before COVID
  filter(yearID >= 1973 & yearID <= 2019)

Batting
```

```{r}
Teams <- Lahman::Teams

Teams <- Teams %>% 
  filter(yearID>=1973 & yearID <= 2019) %>% 
  select(yearID, teamID, park)
```

```{r}
Batting <- Batting %>% 
  right_join(Teams, by = join_by(yearID, teamID))
```

```{r}
Batting %>% 
  right_join(Lahman::People, by = join_by(playerID)) %>% 
  mutate(Age = yearID - birthYear, .after=yearID) %>%
  #Don't consider the people who don't have a birthday in the database
  filter(!is.na(Age))
```

# Dead Ends

```{r}
#Delets the Wrigley field in LA

Parks <- Lahman::Parks %>% 
  filter(park.key != "LOS02")
```

```{r}
#This does not work. Stupid things like "Fenway II" are in there

#TeamID <- 
Teams %>% 
  filter(yearID >= 1973 & yearID <= 2019) %>% 
  select(yearID, teamID, name, park) %>% 
  left_join(Parks, by = join_by(park == park.name)) %>% 
  filter(is.na(state))
```

```{r}
#Manually replace the fields so merge with Parks is possible
Teams <- Lahman::Teams %>% 
  filter(yearID >= 1973 & yearID <= 2019) %>% 
  select(yearID, teamID, name, park) %>% 
  mutate(park = case_when(park=="Fenway Park II" ~ "Fenway Park",
                          park=="Anaheim Stadium" ~ "Angel Stadium of Anaheim",
                          park=="Comiskey Park" ~ "Comiskey Park I",
                          park=="Riverfront Stadium" ~ "Cinergy Field",
                          park=="Royals Stadium" ~ "Kauffman Stadium",
                          park=="Jarry Park" ~ "Parc Jarry",
                          park=="Oakland Coliseum" ~ "Oakland-Alameda County Coliseum",
                          park=="Jack Murphy Stadium" ~ "Qualcomm Stadium",
                          park=="Hubert H Humphrey Metrodome" ~ "Hubert H. Humphrey Metrodome",
                          park=="Exhibition Stadium /Skydome" ~ "Exhibition Stadium",
                          park=="Skydome" ~ "Rogers Centre",
                          park=="Comiskey Park II" ~ "Guaranteed Rate Field",
                          park=="Joe Robbie Stadium" ~ "Sun Life Stadium",
                          park=="Jacobs Field" ~ "Progressive Field",
                          park=="The Ballpark at Arlington" ~ "Rangers Ballpark in Arlington",
                          park=="3Com Park" ~ "Candlestick Park",
                          .default = park))
```

```{r}
Batting %>% 
  right_join(Lahman::Teams %>% filter(yearID>=1973 & yearID <= 2019) %>% select(yearID, teamID, park), by = join_by(yearID, teamID))
```

```{r}
Lahman::People
```