### 2023-11-16 Lahman Analysis/Rough Analysis 2/Analysis3.qmd
Projected a player's trajectory

---
title: "Analysis 3"
format: html
---

# Load Libraries