### 2023-12-11 Lahman Analysis/Formal Analysis/Pres.qmd
Finished first draft of pres

---
title: "Bayesian Data Analysis"
subtitle: "MATH 470 Summary"
author: "Sam Turner"
format: 
  revealjs:
    theme: white
    embed-resources: true
editor: visual
---

## Fellingham and Fisher (2017)

::: {layout-nrow="2"}
![](Figures/FF%20Front.png)

Pros: Bayesian model

-   Bayesian model
:::

## Fellingham and Fisher (2017) Cont.

Cons:

-   Sparingly uses multilevel modeling

    -   Only use multilevel modeling on their orthogonal quartic polynomial

-   Priors are uninformative and don't fit data

    -   Expect average player's HR probability to be 0.0015

    -   Priors effectively suggest that players could have a HR probability between 0 and 1

-   Strange choices of parameters

    -   Age, decade of birth, season of play, home ballpark

## Model

$$
\begin{align}
HR_{nip} &\sim Binomial(AB_{nip},\pi_{nip})\\
\log(\frac{\pi_{nip}}{1-\pi_{nip}}) &= \alpha_{n}+\beta_n\cdot (Age_{ni}-30)+\eta_n\cdot (Age_{ni}-30)^2\\
&+\delta_p+\xi_i
\end{align}
$$

The amount of HR's hit by player $n$ in year $i$ played at park $p$ is binomially distributed, according to player $n$'s AB's and HR probability $\pi$ for year $i$ at park $p$.

HR probability $\pi$ is measured on the logistic scale.

## "Innate" HR Hitting Ability - $\alpha_n$

$$
\begin{align}
\alpha_n&\sim Normal(\mu_0, \sigma_0), n\in\{1,...,657\}\\
\\
\mu_0&\sim Normal(-3.5,0.1)\\
\sigma_0&\sim Exponential(1)
\end{align}
$$

-   Intercept term which represents "innate" hitting ability of players skilled enough to play in MLB

-   -3.5 on the logit scale is ≈0.029 or 2.9%

-   -3.5±0.1 = \[-3.6, -3.4\] = \[0.0266, 0.0323\] = \[2.7%, 3.2%\]

-   -3.5±2 = \[-5.5, -1.5\] = \[0.0041, 0.182\] = \[0.4%, 18.2%\]

## HR Hitting by Age

![](Figures/HRByAgeAll.png){fig-align="center"}

## HR Distribution

![](Figures/pDist.png){fig-align="center"}

## HR Distribution Cont.

![](Figures/pDistSmall.png){fig-align="center"}

## Centered Age Effect - $\beta_n$

$$
\begin{align}
\beta_n &\sim Normal(\mu_1,\sigma_1), n\in\{1,...,657\}\\
\\
\mu_1 &\sim Normal(0,0.1)\\
\sigma_1 &\sim Exponential(10)
\end{align}
$$

-   Multiplicative effect representing how deviation from centered age (Age - 30) affects HR hitting ability

-   Age plays a factor in hitting HR's, but it is likely not very large so we have the priors set near 0 to reflect this

## Centered Age Effect Squared - $\eta_n$

$$
\begin{align}
\eta_n &\sim Normal(\mu_2, \sigma_2),n\in\{1,...,657\}\\
\\
\mu_2 &\sim Normal(0, 0.01)\\
\sigma_2 &\sim Exponential(100)\\
\end{align}
$$

-   Multiplicative effect representing how deviation from centered age squared \[(Age - 30)²\] affects HR hitting ability

-   Used to capture the non-linearity of the data without risk of over-fitting

## HR Hitting by Age

![](Figures/HRByAge20to40.png){fig-align="center"}

## Park Effect - $\delta_p$

$$
\begin{align}
\delta_p &\sim Normal(\mu_5,\sigma_5),p\in\{1,...,88\}\\
\\
\mu_5 &\sim Normal(0,0.01)\\
\sigma_5 &\sim Exponential(10)
\end{align}
$$

-   Intercept term which captures the effect playing in different parks has on HR probabilities

-   Parks differ by both dimensions and altitude which affects HR rates

## Park Overlay

![](Figures/AllPark.png){fig-align="center"}

## Park Overlay Cont.

![](Figures/LongShortPark.png){fig-align="center"}

## Year Effect - $\xi_i$

$$
\begin{align}
\xi_i &\sim Normal(\mu_6,\sigma_6),i\in\{1,...,47\}\\
\\
\mu_6 &\sim Normal(0,0.25)\\
\sigma_6 &\sim Exponential(10)
\end{align}
$$

-   Intercept term which captures the effect playing in different years has on HR probability

-   Changes can occur because of rules, ownership goals, player goals, etc.

-   This term captures those changes without asking why there are changes

## HR's by Year in MLB

![](Figures/HRByYear.png){fig-align="center"}

## HR Proportion by Year in MLB

![](Figures/pByYear.png){fig-align="center"}

## Model (Again)

$$
\begin{align}
HR_{nip} &\sim Binomial(AB_{nip},\pi_{nip})\\
\log(\frac{\pi_{nip}}{1-\pi_{nip}}) &= \alpha_{n}+\beta_n\cdot (Age_{ni}-30)+\eta_n\cdot (Age_{ni}-30)^2\\
&+\delta_p+\xi_i
\end{align}
$$

-   There are $1.17\times10^{12}$ parameters for this model

-   The model predicts that the average player's $\pi$ is about 0.03 (on the normal scale) which is what we observe in the data

-   Uses Bayesian techniques to update estimates based on what the data says - allows for inference

## Robin Yount - Actual $\pi$

![](Figures/Actual651.png){fig-align="center"}

## Robin Yount - Predicted $\pi$

![](Figures/Pred651.png){fig-align="center"}

## Pat Tabler - Actual $\pi$

![](Figures/Actual582.png){fig-align="center"}

## Pat Tabler - Predicted $\pi$

![](Figures/Pred582.png){fig-align="center"}

## Kendrys Morales - Actual $\pi$

![](Figures/Actual420.png){fig-align="center"}

## Kendrys Morales - Predicted $\pi$

![](Figures/Pred420.png){fig-align="center"}

## Goodness of Fit - Trace-plots

![](Figures/alphaTrace.png){fig-align="center"}

## Conclusions

-   Will our model make the Hood math department excellent gamblers?

    -   No. But it did a fine job at adapting predictions of players and being "right" on average

-   Areas of future research

    -   Player archetypes and physical characteristics

    -   Considering more for a longer time interval (like Fellingham and Fisher (2017))

    -   Better data (advanced metrics or play-by-plays)

## Conclusions Cont.

-   We believe we did better than Fellingham and Fisher (2017), but without access to their model and computational resources it is difficult to determine